---
title: "CHL8010: Statistical Programming and Computation in Health Data"
author: "Week 4 In-class Assignment"
date: "2024-09-30"
format: pdf
---

```{r, warning=FALSE}
library(here)
library(ggcorrplot)
library(ggplot2)
library(dplyr)
source(here("R", "merged_dataset.R"))
```

## Perfect your GitHub repo

Some of you may still need to organize your GitHub repo. Use this time to do that. When you are confident with your repo, let me know -- I will try to reproduce your code.

Your final data should have the following variables (you might have slightly different variable names).

```{r}
finaldata <- read.csv(here("data", "analytical", "finaldata.csv"), header = TRUE)
names(finaldata)
```

Observations from Canada should look like this...

```{r}
finaldata %>%
  dplyr::filter(country_name == "Canada")
```

Observations from Ecuador should look like this...

```{r}
finaldata %>%
  dplyr::filter(country_name == "Ecuador")
```

## Exploratory data analysis

Use the rest of the class time to explore the final data that will be used for analysis starting next week. At the end of the class, write a summary of your findings and push your **Quarto document (pdf)** to your repo.

```{r}
summary(finaldata)
```

```{r}
# Check for missing values
colSums(is.na(finaldata))
```
gdp1000, popdens, urban, male_edu, temp, rainfall1000, MarMor, InfMort, NeonatalMort 
all have missing data 

MarMor has a lot of missing data



```{r}
#Bivariate analysis

#GDP and Population density
ggplot(finaldata, aes(x=gdp1000, y=popdens)) + 
  geom_point() + 
  labs(title="GDP vs Population Density")

#note missing values were removed
```



```{r}

#Bivariate analysis

#temp vs. rainfall1000
ggplot(finaldata, aes(x=temp, y=rainfall1000)) + 
  geom_point() + 
  labs(title="Temperature vs Rainfall")
```
Lower temperature countries have less rainfall whereas for higher temperature, it varies

```{r, warning=FALSE}

#Bivariate analysis

#gdp1000 vs. Under5Mort
ggplot(finaldata, aes(x=gdp1000, y=Under5Mort)) + 
  geom_point() + 
  labs(title="GDP vs Under 5 mortality")
```
Countries with low GDP have substantially greater mortality rates for children under 5. 

```{r, warning = FALSE}

#Bivariate analysis

#gdp1000 vs. NeonatalMort
ggplot(finaldata, aes(x=gdp1000, y=NeonatalMort)) + 
  geom_point() + 
  labs(title="GDP vs Neonatal Mortality")
```
Countries with lower GDP have substantially higher neonatal mortality rates

```{r, warning=FALSE}

#Bivariate analysis

#gdp1000 vs. InfMort
ggplot(finaldata, aes(x=gdp1000, y=InfMort)) + 
  geom_point() + 
  labs(title="GDP vs Infant Mortality")
```

Countries with lower GDP have substantially higher infant mortality rates

